### 2021-08-07 curriculum_vitae/resume.tex
started resume

\documentclass{article}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}

% Set page geometry
\usepackage[margin=1.5cm,headsep=0.5cm]{geometry}

\input{../preamblex}

\begin{document}

\begin{minipage}{0.63\textwidth}
\begin{flushleft}
\textbf{\textsc{\Huge{Nicholas D. Kostin}}}
\end{flushleft}
\end{minipage}
~
\begin{minipage}{0.35\textwidth}
\begin{flushright}
\parbox[c]{\textwidth}{Website: \url{https://nkostin.xyz} \\ Email: \texttt{contact@nkostin.xyz} \\ GitHub: \url{https://github.com/nkostin4} \\ Mobile: +1 719-271-9108}
\end{flushright}
\end{minipage}

\vspace{1em}

\textbf{\Large{Education}} \\[-0.5em]
\rule{\textwidth}{1pt}

\hspace{1.5em} \textbf{Colorado School of Mines} \hfill Golden, CO

\hspace{1.5em} Major: \textbf{Engineering Physics} \hfill BS May 2022

\hspace{1.5em} Minor: \textbf{Computational and Applied Mathematics}

\vspace{1em}

\hspace{1.5em} Major GPA: \textbf{4.00} / Cumulative GPA: \textbf{4.00}

\vspace{1em}

\textbf{\Large{Work Experience}} \\[-0.5em]
\rule{\textwidth}{1pt}

\hspace{1.5em} \textbf{Teaching Assistant} \hfill Golden, CO

\hspace{1.5em} \textit{Physics II --- Electromagnetism and Optics} \hfill January 2019 --- May 2021

\end{document}