### 2021-09-06 report/chapters/opening.tex
more memes and finished serious changes

\documentclass[class=report,crop=false]{standalone}
\usepackage[margin=2.5cm,headsep=0.5cm]{geometry}
\input{../../preamblex}

\begin{document}

\section{An Opening Section}

\blindtext[5]

\begin{equation}
    x = 63
\end{equation}

\blindtext[1]

\begin{equation}
    z = 6
\end{equation}

\blindtext[1]

\begin{equation}
    w = 1986
\end{equation}

\blindtext[1]

\section{Continuing Forward}

\blindtext[5]

\begin{equation}
    y = 4
\end{equation}

\blindtext[1]

\begin{equation}
    a = 63
\end{equation}

\blindtext[1]

\begin{figure}[H]
\captionsetup{width=0.8\textwidth,labelfont={color=BLACK,bf},textfont={color=BLACK}}
\centering
\begin{tikzpicture}[scale=0.5]
    \fill[PURPLE] (0,0) circle (4);
\end{tikzpicture}
\caption{Here's an example figure.}
\end{figure}

\blindtext[1]

\begin{figure}[H]
\captionsetup{width=0.8\textwidth,labelfont={color=BLACK,bf},textfont={color=BLACK}}
\centering
\begin{tikzpicture}[scale=0.5]
    \fill[GREEN] (0,0) circle (4);
\end{tikzpicture}
\caption{Look at this! Another example figure!}
\end{figure}

\blindtext[1]

\begin{figure}[H]
\captionsetup{width=0.8\textwidth,labelfont={color=BLACK,bf},textfont={color=BLACK}}
\centering
\begin{tikzpicture}[scale=0.5]
    \fill[ORANGE] (0,0) circle (4);
\end{tikzpicture}
\caption{A little hard to believe. Another example figure?}
\end{figure}

\blindtext[1]

\begin{equation}
    c = 660
\end{equation}

\blindtext[1]

\begin{equation}
    t = 4
\end{equation}

\blindtext[1]

\begin{equation}
    d = 175
\end{equation}

\blindtext[5]

\section{More Explanation}

\blindtext[5]

\begin{figure}[H]
\captionsetup{width=0.8\textwidth,labelfont={color=BLACK,bf},textfont={color=BLACK}}
\centering
\begin{tikzpicture}[scale=0.5]
    \fill[BLUE] (0,0) circle (4);
\end{tikzpicture}
\caption{Is it true? Another example figure?}
\end{figure}

\blindtext[5]

\begin{figure}[H]
\captionsetup{width=0.8\textwidth,labelfont={color=BLACK,bf},textfont={color=BLACK}}
\centering
\begin{tikzpicture}[scale=0.5]
    \fill[RED] (0,0) circle (4);
\end{tikzpicture}
\caption{It looks to be another example figure.}
\end{figure}

\blindtext[5]

\begin{figure}[H]
\captionsetup{width=0.8\textwidth,labelfont={color=BLACK,bf},textfont={color=BLACK}}
\centering
\begin{tikzpicture}[scale=0.5]
    \fill[GOLDE] (0,0) circle (4);
\end{tikzpicture}
\caption{This has to be the last figure.}
\end{figure}

\section{Mobility}

\blindtext[5]

\begin{equation}
    f = 002
\end{equation}

\blindtext[1]

\begin{equation}
    g = 10
\end{equation}

\blindtext[5]

\subsection{Stability}

\blindtext[5]

\begin{equation}
    h = 5
\end{equation}

\blindtext[1]

\begin{equation}
    k = 20
\end{equation}

\blindtext[5]

\section{Closing}

\blindtext[5]

\begin{equation}
    h = 5
\end{equation}

\blindtext[1]

\begin{equation}
    k = 20
\end{equation}

\blindtext[1]

\begin{equation}
    u = 3
\end{equation}

\blindtext[1]

\begin{equation}
    t = 3
\end{equation}

\blindtext[10]

\end{document}